Ae. albopictus CPP manuscript

### 2020-02-05 MS.Rmd
Removed redundant figures

```{r filter papers, echo=FALSE, include=FALSE}
#removing papers with <60% or >100% initial count
blockAB_unfiltered<-blockAB
blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
#test removing papers that had premature larval hatch on paper
#badpapers<-c("333-1", "334-1", "335-1", "336-1", "337-1", "338-1", "339-1", "340-1", "341-1", "342-1", "343-1","366-1", "367-1", "368-1", "369-1", "370-1", "371-1", "372-1", "373-1", "374-1", "375-1", "376-1")
#blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) %>% filter(!PaperID %in% badpapers)
```

@@

```{r incidence_lat, echo=FALSE, include=FALSE}
DI_short<-filter(DI_summary, pp==8)
ggplot(DI_short, aes(x=lat, y=DI, color=country)) + geom_point() + theme_bw() +stat_smooth()
```

@@

ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
  scale_color_manual(values=c("black","gray"), name="Country") +
  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + xlab("Latitude") + ylab("Critical photoperiod (h)")

ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country)) + 
geom_point(size=3, shape=21) + scale_fill_manual(values=c("black", "gray"), name="Country") + stat_smooth(method=lm, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
  facet_grid(~year) +
  guides(fill=guide_legend(override.aes=list(shape=21))) +
  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + 
  xlab("Latitude") + ylab("Critical photoperiod (h)")

```{r incidence_lat_year, echo=FALSE, include=FALSE}
DI_short_2008<-filter(DI_summary_2008, photoperiod==8)
DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short_2<-rbind(DI_short_2008, DI_short_2018)
write_csv(DI_short_2, "diapause_incidence_8h_2008_2018.csv")
incidence_lat_year_plot<-ggplot(DI_short_2, aes(x=lat, y=DI, color=year, label=pop)) + scale_color_manual(values=c("gray","black")) + 
  geom_point() + facet_grid(~country) 
country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm"))
country_DI<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position=c(0.9,0.1)) + facet_grid(~year)
```

@@

```{r day_function, echo=FALSE, include=FALSE}
findDate<-function(lat, pp, year) {
#generate day lengths at latitude between July 1 and Dec 31
x<-daylength(lat,182:365)
#find position of day length in the daylength vector that is closest to photoperiod queried
daynumber<-which.min(abs(x-pp))
#convert day number to date
origin<-as.character(paste(year,"-07-01",sep=""))
return(as.character(as.Date(daynumber, origin=origin)))
}
```

```{r approx_dates, include=FALSE, echo=FALSE, include=FALSE}
cpp_2$date<-mapply(findDate, lat=cpp_2$lat, pp=cpp_2$Estimate, year=cpp_2$year)
cpp_2$estimated_date<-format(as.Date(cpp_2$date), "%mm-%dd")
cpp_2$estimated_date<-as.Date(cpp_2$estimated_date, format="%mm-%dd")
date_plot<-ggplot(cpp_2, aes(x=lat, y=estimated_date, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") +  facet_grid(~country) + theme_bw() + stat_smooth(method=lm, se=FALSE)
```

@@

plot_grid(us_map_bw, japan_map_bw)

@@

```{r fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
incidence_lat_year_plot
```

@@

ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")

ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray")) + 
scale_shape_manual(values=c(21,24)) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_color_manual(values=c("black","gray")) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")

@@

```{r fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
cpp_08_18_shifts
```

@@

```{r fig.cap="Figure S1: Discrepancies between initial egg counts and final calculated egg counts on oviposition papers.", echo=FALSE}
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point() + xlab("Number of viable eggs on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw()
```

```{r fig.cap="Figure S2. Critical photoperiod calculated with and without filtering oviposition papers based on criteria described. Line represents x=y. Outlier point corresponds to population PBS.", echo=FALSE}
#comparing CPP values calculated with and without filtering out unreliable oviposition papers
cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% filter(year==2018)
cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% filter(year==2018)
names(cpp_filtered)[5]<-"CPP_est_filt"
names(cpp_unfiltered)[5]<-"CPP_est_unfilt"
cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
```

```{r fig.cap="Figure S3: Photoperiod response curves", echo=FALSE}
DI_plot 
```

```{r fig.cap="Supplemental figure TBD. Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
DI_pies_2018
```

@@

```{r}
#Dose response curve plot from Jen
#cpp<-read.table("CPP.csv", header=TRUE, sep=",")
#cpp

#fm1<-cbind(hatched, nothatched)~photoperiod
#cppmodel<-glm(fm1, data=cpp, family=binomial())

#summary(cppmodel)

#fitteddata1<-predict(cppmodel,type="response")
#fitteddata1

#plot(cpp$photoperiod, fitteddata1, ylab="Probability of hatching",
 #    xlab="photoperiod", ylim=c(0,1))
  #   lines(cpp$photoperiod,fitteddata1, lty=1)
#y<-cpp$hatched/(cpp$hatched+cpp$nothatched)
#text(cpp$photoperiod,y)
#hours<-(cppmodel$coefficients["(Intercept)"])/cppmodel$coefficients["photoperiod"]
#hours

>>>>

```{r filter papers, echo=FALSE, include=FALSE}
#removing papers with <60% or >100% initial count
blockAB_unfiltered<-blockAB
blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
```

@@

@@

cpp_08_18_shifts_bw<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
  scale_color_manual(values=c("black","gray"), name="Country") +
  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + xlab("Latitude") + ylab("Critical photoperiod (h)")
```

```{r incidence_lat_year, echo=FALSE, include=FALSE}
DI_short_2008<-filter(DI_summary_2008, photoperiod==8)
DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short_2<-rbind(DI_short_2008, DI_short_2018)
write_csv(DI_short_2, "diapause_incidence_8h_2008_2018.csv")
```

@@

@@

map_panels_bw<-plot_grid(japan_map_bw, us_map_bw)

@@

@@

@@

```{r fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
cpp_08_18_shifts_bw
```

@@

```{r, echo=FALSE}
yield_plot<-ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point(alpha=0.5) + xlab("Number of viable eggs on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw()

#comparing CPP values calculated with and without filtering out unreliable oviposition papers
cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% filter(year==2018)
cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% filter(year==2018)
names(cpp_filtered)[5]<-"CPP_est_filt"
names(cpp_unfiltered)[5]<-"CPP_est_unfilt"
cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
filter_compare_plot<-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
```

```{r, echo=FALSE}
quality_control_fig<-plot_grid(yield_plot, filter_compare_plot, labels="AUTO")
quality_control_fig
```

```{r, echo=FALSE}
DI_plot 
```

@@
